---
title: "Simulations Pre-Class Project"
date: "Due March 13, 2017 at 5:00pm"
output:
  html_document


---

<style type="text/css">
.table {

    width: 80%;
    margin-left:10%; 
    margin-right:10%;
}
</style>
```{r,setup, echo=FALSE, cache=TRUE}
## numbers >= 10^5 will be denoted in scientific notation,
## and rounded to 2 digits
options(scipen = 3, digits = 3)
```




#Project Goals:


With this project we will simulate a famoues probability problem. This will not require knowledge of probability or statistics but only the logic to follow the steps in order to simulate this problem. This is one way to solve problems by using the computer. 

 1. **Gambler's Ruin**: Suppose you have a bankroll of $1000 and make bets of $100 on a fair game. By simulating the outcome directly for at most 5000 iterations of the game (or hands), estimate:
    a. the probability that you have "busted" (lost all your money) by the time you have placed your one hundredth bet. 
```{r}
gamblers = function(bankroll, num.bets){
  #place n bets
  outcome = rbinom(n = num.bets, size = 1, prob = 0.5)
  for (j in 1:length(outcome)){
    if(outcome[j] == 1){
    #win
    bankroll = bankroll + 100
    } else {
    #lose
    bankroll = bankroll - 100
    }
    if(bankroll == 0){
      break
    }
  }
  return(bankroll)
}

#create 5000 iterations of the game
profits = replicate(5000, gamblers(1000, 100))

#find probability that you have busted
prob_bust = length(profits[profits == 0]) / length(profits)
prob_bust
```
    b. the probability that you have busted by the time you have placed your five hundredth bet by simulating the outcome directly. 
```{r}
#create 5000 iterations of the game
profits_500 = replicate(5000, gamblers(1000, 500))

#find probability that you have busted
prob_bust = length(profits_500[profits_500 == 0]) / length(profits_500)
prob_bust
```
    c. the mean time you go bust, given that you go bust within the first 5000 hands.
```{r}
time_bust = c()

for(i in 1:5000){
  #start with $1000
  bankroll = 1000
  #place 100 bets
  outcome = rbinom(n = 5000, size = 1, prob = 0.5)
  for (j in 1:length(outcome)){
    if(outcome[j] == 1){
      #win
      bankroll = bankroll + 100
    } else {
      #lose
      bankroll = bankroll - 100
    }
    if(bankroll == 0){
      #record the time when you go bust
      time_bust[i] = j
      break()
    }
  }
}

mean_bust = mean(time_bust, na.rm = TRUE)
mean_bust

```
    d. the mean and variance of your bankroll after 100 hands (including busts).
```{r}
#use profits vector from part a
mean(profits)
var(profits)
```
    e. the mean and variance of your bankroll after 500 hands (including busts).
```{r}
#use profits_500 vector from part a
mean(profits_500)
var(profits_500)
```
 
Note: you *must* stop playing if your player has gone bust. How will you handle this in the `for` loop?

2. **Markov Chains**. Suppose you have a game where the probability of winning on your first hand is 48%; each time you win, that probability goes up by one percentage point for the next game (to a maximum of 100%, where it must stay), and each time you lose, it goes back down to 48%. Assume you cannot go bust and that the size of your wager is a constant $100.
    a. Is this a fair game? Simulate one hundred thousand sequential hands to determine the size of your return. Then repeat this simulation 99 more times to get a range of values to calculate the expectation.
```{r}
set.seed(123)

markov = function(bankroll, num.hands, prob_start){
  prob = prob_start
  #iterate through number of hands
  for (i in 1:num.hands){
    if_win = rbinom(n = 1, size = 1, prob = prob)
    #if you win, add 0.01 to the probability
    if (if_win == TRUE) {
      if(prob < 1){
        prob = prob + 0.01
        }
      bankroll = bankroll + 100
    }
    #if you lose, go back to original probability
    if (if_win == FALSE){
      prob = prob_start
      bankroll = bankroll - 100
    }
  }
  return(bankroll)
}

#simulate game 100 times
sim = replicate(100, markov(0, 10000, 0.48))

#find expected return
exp_return = mean(sim)
exp_return
```
This is not a fair game since the expected value is not 0
    b. Repeat this process but change the starting probability to a new value within 2% either way. Get the expected return after 100 repetitions. Keep exploring until you have a return value that is as fair as you can make it. Can you do this automatically?
```{r, cache = TRUE}
#start from 0.48 since we know that we are losing at p = 0.48

#create a vector of probabilities to use in the function
values = seq(0.48, 0.5, by = 0.001)

#this for loop evaluates markov using different starting probabilities
for (j in values){
  exp_return = mean(replicate(100, markov(0, 100000, j)))
  #check if the expected return is close to 0
  if(exp_return < 2500 && exp_return > -2500){
    print(exp_return)
    print(j)
    break
  }
}
```
The game is fair when p = 0.489 since at this starting probability the expected return is close to 0.
    c. Repeat again, keeping the initial probability at 48%, but this time change the probability increment to a value different from 1%. Get the expected return after 100 repetitions. Keep changing this value until you have a return value that is as fair as you can make it. 
```{r}
#increase in prob 0.0125
markov = function(bankroll, num.hands, prob_start, j){
  prob = prob_start
  for (i in 1:num.hands){
    if_win = rbinom(n = 1, size = 1, prob = prob)
    if (if_win == TRUE) {
      if(prob < 1){
        prob = prob + j
        }
      bankroll = bankroll + 100
    }
    if (if_win == FALSE){
      prob = prob_start
      bankroll = bankroll - 100
    }
  }
  return(bankroll)
}

for (k in seq(0.01, 0.02, by = 0.001)){
  mean(replicate(100, markov(0, 100000, 0.48, k)))
  if(exp_return < 2500 && exp_return > -2500){
  print(exp_return)
  print(k)
  break
  }
}
```


