# pre-class


Make sure you commit this often with meaningful messages. 




### Question 1:

Using a loop, print the integers from 1 to 50. 

```{r}
for(i in 1:50){
  print(i)
}
```

### Question 2:

A.  Using a loop, add all the integers between 0 and 1000.
```{r}
#int contains integers from 0 to 1000
int = seq(0, 1000, by = 1)
  
#initialize tot_sum to 0
tot_sum = 0

#this for loop iterates through each element in int
for(i in int){
  tot_sum = tot_sum + i
}
```

B. Now, add all the EVEN integers between 0 and 1000 (hint: use seq())
```{r}
#even_int contains even integers between 0 and 1000
even_int = seq(0, 1000, by = 2)

#initialize sum_even to 0
sum_even = 0

#this for loop iterates through each element in even_int 
for (i in even_int){
  sum_even = sum_even + i
}
```

C. Now, repeat A and B WITHOUT using a loop.
```{r}
#repeat part a
sum(int)

#repeat part b
sum(even_int)
```

### Question 3:

Here is a dataframe of survey data containing 5 questions :

```{r, eval=FALSE}
survey <- data.frame(
                     "participant" = c(1, 2, 3, 4, 5, 6),
                     "q1" = c(5, 3, 2, 7, 11, 0),
                     "q2" = c(4, 2, 2, 5, -10, 99),
                     "q3" = c(-4, -3, 4, 2, 9, 10),
                     "q4" = c(-30, 5, 2, 23, 4, 2),
                     "q5" = c(88, 4, -20, 2, 4, 2)
                     )
```
The response to each question should be an integer between 1 and 5. Obviously, we have some bad values in the dataframe. The goal of this problem is to fix them.

A. Using a loop, create a new dataframe called survey.clean where all the invalid values (those that are not integers between 1 and 5) are set to NA.
```{r}
survey.clean = survey

#this for loop iterates through each row and each column of survey
for(i in 1:nrow(survey)){
  for(j in 2:ncol(survey)){
    #check if the i, j entry in survey is less than 1 or greater than 5
    if(survey[i, j] < 1 | survey[i,j] > 5){
      #change the i, j entry in survey.clean to NA
      survey.clean[i, j] = NA
    }
  }
}
```

B. Now, again using a loop, add a new column to the dataframe called “invalid.answers” that indicates, for each participant, how many bad answers they gave.
```{r}
#initialize num_na to 0
num_na = 0

#this for loop iterates through each row and column of survey.clean
for(i in 1:nrow(survey.clean)){
  for(j in 2:ncol(survey.clean)){
    #check if the entry is NA
    if(is.na(survey.clean[i, j])){
      #add 1 to the current number of NA's
      num_na = num_na + 1
    }
  }
  #create a new column with the num_na value
  survey.clean$invalid.answers[i] = num_na
  #reset the sum to 0 before you look at the next row
  num_na = 0
}
```