# pre-class


Make sure you commit this often with meaningful messages. 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.

```{r}
f1 <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}

f2 <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}

f3 <- function(x, y) {
  rep(y, length.out = length(x))
}
```
The first function takes a string and checks whether that string starts with a specified prefix, so it could be called check_prefix.

The second function takes a vector and removes the last element, so it could be called remove_last.

The third function takes in two vectors and replicates the second one until it has the same number of elements as the first one, so it could be called replicate_vec.

2. Compare and contrast rnorm() and MASS::mvrnorm(). How could you make them more consistent?  
rnorm is used to generate a random normally distributed sample. It takes 3 arguments: n, mean, and sd. MASS::mvrnorm produces one or more samples from a multivariate normal distribution and it also takes 3 arguments: n, mu and sigma. To make them more consistent we should use the same names for the arguments. 

3. Use `lapply()` and an anonymous function to find the coefficient of variation (the standard deviation divided by the mean) for all columns in the mtcars dataset. 

```{r}
#this function calculates the coefficient of variation
coeff_of_var = function(vec){
  sd(vec) / mean(vec)
}

lapply(mtcars, FUN = coeff_of_var)
```

4. Use vapply() to:
    a. Compute the standard deviation of every column in a numeric data frame.
```{r}
vapply(mtcars, FUN = sd, numeric(1))
```
    b. Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you’ll need to use vapply() twice.)
```{r}
#iris[vapply(iris, is.numeric, logical(1))] selects only the numeric columns
vapply(iris[vapply(iris, is.numeric, logical(1))], FUN = sd, numeric(1))
```
